### 2023-11-27 thesis/chapters/02-analysis-C-diagnostics.tex
Incremental learning in thesis

\subsection{Concept drift}
% TODO: ADWIN (ADaptive WINdowing)
% Kolmogorov-Smirnov Windowing method for concept drift detection.

@@

@@

\subsection{Classification}
% TODO: HoeffdingTreeClassifier
% https://analyticsindiamag.com/a-beginners-guide-to-hoeffding-tree-with-python-implementation/
% https://www.cms.waikato.ac.nz/~abifet/book/chapter_5.html

Accurate multi-class classification of machine fault causes out of the characteristics of known ones is a much more difficult task than novelty detection. Universal enough fault baseline has to be recorded and transformed into feature space. Interactions among fault manifestations have to be accounted for. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem not production ready. Therefore, we opt to use a more rudimentary model.
\bigbreak

\textbf{K-nearest neighbors} (kNN) assigns the data point to the class where the majority of $k$ closest instances belong (Fig.~\ref{fig:knn}). This means it can work in the semi-supervised environment because it can infer labels just from knowing a few annotations.

@@

% TODO: Spatial database
% TODO: Sliding WIndow-based Nearest Neighbor (SWINN)

The sense of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, or \emph{Manhatann distance} (Tab.~\ref{tab:knn-distance})~\cite{sheng_review_2020}. The demanding neighborhood queries are sped up by a Metric search tree data structure such as \emph{M-Tree} (Fig.~\ref{fig:m-tree}). The optimal $k$ parameter is set in supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate.

@@

\subsection{Online model evaluation}
% TODO

>>>>

@@

Quality of clustering results is evaluated by \emph{Silhouette score} in the range [-1; 1]. It demands points within clusters to have high cohesion and at the same time to have large separation from other clusters. The low score indicates a too little or too large number of clusters~\cite{rousseeuw_rousseeuw_1987}.

@@

\subsection{Classification}
Accurate multi-class classification of machine fault causes out of the characteristics of known ones is a much more difficult task than novelty detection. Universal enough fault baseline has to be recorded and transformed into feature space. Interactions among fault root causes have to be considered. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem not production-ready and hard to explain. Therefore, we opt to use a simpler model.

Performance of classification is estimated by several metrics on the validation set obtained using hold-out or cross-validation techniques. Frequently used quantities for classifier model evaluation include accuracy, precision, recall, f1 score, area under ROC curve, and counts of hits and misses in confusion matrix.
\bigbreak

\textbf{K-nearest neighbors} (kNN) assigns the data point to the class where the majority of $k$ closest instances belong (Fig.~\ref{fig:knn}). It means it can work in the semi-supervised environment because it can infer labels just from knowing a few annotations. The major drawback of kNN is a preference for the majority class in imbalanced class sizes datasets~\cite{shi_improving_2020}. The issue is mitigated with class weights or resampling classes by oversampling or undersampling. kNN algorithm requires features to be normalized to assign the same importance to each predictor.

@@

The sense of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, or \emph{Manhatann distance} (Tab.~\ref{tab:knn-distance})~\cite{sheng_review_2020}. The optimal $k$ parameter is set in supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate. The demanding neighborhood queries are sped up using spatial index in spatial databases that utilizes search tree such as kd-tree, R-tree, or M-tree.

@@

\subsection{Incremental learning}
Online and incremental machine learning operates in the streaming setting updating the model parameters with each new incoming event or in mini-batches. This approach finds its use in big data processing when the whole dataset is not available in advance or cannot be processed at once because of memory limitations.

There are some additional obstacles to watch out for with incremental learning in comparison to batch learning~\cite{gepperth_incremental_2016}:
\begin{enumerate}
    \itemsep0pt
    \item \textbf{Concept drift} is defined as the change in data distribution function over time. The two types of concept drift are virtual and real. In virtual drift, changes occur only in the input distribution. Real drift means that the alteration comes to underlying functionality. Concept shift occurs with the abrupt change.
    \item \textbf{Stability-plasticity tradeoff} concerns the speed with which the model adapts to new information. The model can react quickly making it less stable, or retain patterns for longer but become irresponsive to sudden shifts.
    \item \textbf{Model complexity} should be adjustable to ensure flexibility in unforeseen circumstances. Simpler models in the ensemble can also further increase prediction robustness. Resource limitation bound complexity from above.
    \item \textbf{Memory model} can store aggregates from seen observations and its typical examples or finite window of latest samples with forgetting factor.
\end{enumerate}

\textbf{Model benchmarking} in incremental learning is achieved by comparing models to their batch counterparts or using progressive validation~\cite{blum_beating_1999, halford_correct_2020}. It has been shown that incremental clustering algorithms have overall worse accuracy than batch versions~\cite{gepperth_incremental_2016}. In the validation process, precautions should be taken to prevent data leakage from future events into the past.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{assets/incremental-learning.png}
    \caption{Incremental learning deployment architecture~\cite{gaia_online_2022}}
    \label{fig:online-learn-arch}
\end{figure}

Deployment of an online machine learning model alongside supporting services is different than in established MLOps processes. Model store and Inference service components are supplemented with Labelling and Learning service~\cite{gaia_online_2022}. Their goal is to tune model parameters gradually as additional ground truth labels are provided. Labels can be provided later in a scheme called ``log and wait'', but data features are stored until such time.

### 2023-11-27 thesis/chapters/03-design.tex
Incremental learning in thesis

\section{Preparatory exploration}
The wood processing factory we plan to collaborate with has to choose machines that are worthy of condition monitoring. In the meantime, the data processing pipeline will be put together in Python language and packages for scientific computing. The pipeline will consist of the following stages.

\begin{enumerate}
    \itemsep0pt
    \item Detrending
    \item Adaptive noise cancellation of the background interference
    \item Vector of all features
    \begin{itemize}
        \itemsep0pt
        \item Time domain: statistical measures (Tab.~\ref{tab:td-features})
        \item Frequency domain: PSD estimation with FFT and Welch averaging with the resolution of 1 Hz combined with Hann window (Tab.~\ref{tab:fd-features})
        \item Time-frequency domain: energy and entropy in wavelet coefficients from WPD and EWT filter banks.
    \end{itemize}
    \item Feature selection on evaluation datasets and experimental measurements from the factory to prune away irrelevant features with pearson correlation, Fisher score, and mutal information.
    \item Model evaluation and comparison of novelty detection methods and precision of classification with different sets of predictors. The range of optimal parameters will be found for the DenStream ($\mu$, $\epsilon$, $beta$, $\lambda$), Half-Space Trees (window size, ensemble size), and kNN (distance metric, $k$ neighbours). Evaluation metrics associated with confusion matrices will be used like accuracy, precision, true positive rate, and false positive rate.
\end{enumerate}

@@

\section{Vibration processing}
 \begin{itemize}
 \itemsep0pt
\item \textbf{Input:} Samples from three-axis MEMS accelerometers, RPM tachometer, Noise background
\item \textbf{Output:} machine health status / type of fault
\item \textbf{Output on demand}: Control chart of trend features
\end{itemize}

>>>>

\section{Machine learning pipeline}

Mafaulda dataset
Jupyter notebook - Python
\begin{enumerate}
    \itemsep0pt
    \item Detrending
    \item (Optional) Adaptive noise cancellation of the background interference
    \item Vector of all features
    \begin{itemize}
        \itemsep0pt
        \item Time domain: statistical measures (Tab.~\ref{tab:td-features})
        \item Frequency domain: PSD estimation with FFT and Welch averaging with the resolution of 1 Hz combined with Hann window (Tab.~\ref{tab:fd-features})
    \end{itemize}
    \item Feature selection on evaluation datasets and experimental measurements from the factory to prune away irrelevant features with pearson correlation, Fisher score, and mutal information.
    \item Model evaluation and comparison of novelty detection methods and precision of classification with different sets of predictors. The range of optimal parameters will be found for the DenStream ($\mu$, $\epsilon$, $beta$, $\lambda$), Half-Space Trees (window size, ensemble size), and kNN (distance metric, $k$ neighbours). Evaluation metrics associated with confusion matrices will be used like accuracy, precision, true positive rate, and false positive rate.
\end{enumerate}

@@

\section{Research questions}
Main goals:
	- How does vibration data look like (spectral and time domain)?
	- Create dataset of machinery vibrations
	- Reduce number of features (samples) send from edge device - to minimal amount
	- Adapt to features changing over time (RiverML can handle it, find proof)
	- Evaluate model performance on all features and different feature sets
	- Keep in mind: data ordering!

\section{Data preparation}
- Mafaulda dataset
	- Choose 4 faults - normal (N), imbalance (I), vertical (VM) and horizonal misalignment (HM)
	- Convert severity to 0 - 1 scale.
		- Label samples above 0.5, 0.7, 0.9 as anomalous
	- Choose rpm range 2900 $\pm$ 500 (around rotational speed of machinery with some spread, some features are correlated with rpm 
	- Counts before split: 
		- Split to 5 parts of 1 second intervals
	- Two measurement points A, B
	- Detrending by removing mean
	- Calculate features: temporal, spectral (in 5 window sizes): welch averaging, hann window

	- Batch: 
		- Balancing dataset with oversampling minority classes	
		- Hold-out validation - split to train and test set
	- Online: Order by severity

- Measurement of custom datasets 
	- one 3-axis accelerometer on top of scroll compressor of air conditioning
		- one in the base to capture noise
	- water pump

	- columns: t, x, y, z: rpm
	- 20 s recordings - split to 5 parts - extract features
	- do PCA to explore see differences
	- label as worse, better (classes are hard to provide)

\section{Feature relevance}
- choose 3 features
- supervised
	- highly correlated features >= 0.9:
		- Temporal domain (A)
			- std, rms, 1.000000
			- margin, impulse, 0.999079
			- crest, impulse, 0.996807
			- crest, margin,	0.992548
			- pp, max, 0.970055
			- max, (std, rms), 0.938198
			- pp, (std, rms), 0.938180 / 0.928
		- Spectral domain (A)
			- same feature in different window sizes (0.99)
			- kurtosis, skewness, 0.980618
			- std, energy, 0.960110
			- entropy, noisiness, 0.945514

	- calcultate euclidian measure of feature 
	- correlation point biserial, fisher, mutual information
		- Sperman rank correlation: corr, f_stat, 1.000000
	- calculate feature rank
		- Temporal (A) - Remove corr: pp, skewness, margin
			pp	1.442250
			std	2.080084
			max	2.714418
			rms	3.174802
			skewness	6.082202
			margin	6.868285

			(B)			
			skewness	1.817121
			rms	1.817121
			std	2.289428
			kurtosis	4.308869
		- Spectral (A) - multiple window sizes 
			- (energy, roll off, std)

- online learning features
	- evolution of feature importace

\section{Nearest neighbour classifier}
- kNN (fault and anomaly)
	- Offine (baseline)
	- Online (evolution)
- parameters: k-neigh, distance
- metrics: accuracy, precision, recall

\section{Clustering}
- Density based clustering
	- DBSCAN (all features, PCA, subset of best)
	- DenStream (all features, subset of best)

\section{Novelty detection}
- Offline (Isolation forest)
- Online (Half-space Tree)

\section{Infrastructure deployment}
Future work: engineer new feature, deploy, tweek models, validate on real dataset

 \begin{itemize}
 \itemsep0pt
\item \textbf{Input:} Samples from three-axis MEMS accelerometers, RPM tachometer, Noise background
\item \textbf{Output:} machine health status / type of fault
\item \textbf{Output on demand}: Control chart of trend features
\end{itemize}
